Fractional Types

### 2016-12-19 typespp.tex
simplify diagrams (by omitting, like in the rest of the circuits, applications of id combinators).  Add more words to their description.  These are now quite simple.

ccc have a notion of higher-order functions, trace, etc. Talk about these
examples.

\begin{lemma}
  \label{lem:cong}
  For $\mathbb{v} : \mathbb{T}_1 \circledast \mathbb{T}_2$,
  $p : \mathbb{T}_1 \leftrightblackspoon \mathbb{T}$, we have
  $\texttt{cong}^{l}_{p}(\mathbb{v}) : \mathbb{T} \circledast \mathbb{T}_2$, and
  for $q : \mathbb{T}_2 \leftrightblackspoon \mathbb{T}$, we have
  $\texttt{cong}^{r}_{q}(\mathbb{v}) : \mathbb{T}_1 \circledast \mathbb{T}$.
  \begin{proof}
    It follows from the fact that $\circledast$ is an endo-bi-functor.
  \end{proof}
\end{lemma}

@@

Given a \textrm{/-combinator} $p : \#c \leftrightblackspoon \#d$ which is a
morphism in the compact closed category, we have a dual morphism
$p^{\ast} : 1/\#d \leftrightblackspoon 1/\#c$. This is shown by building a
reversible circuit in figure~\ref{fig:dual}.

\begin{figure}[htb]
  \centering
  \begin{tikzpicture}[scale=0.9,every node/.style={scale=0.7},highlight/.style={line width=4pt,orange!70}]
    \path (-1,1) edge node[above] {\footnotesize{$1/{\textsf{\#d}}$}} (0,1);

@@

    \path (1,1.75) edge node[above] {\footnotesize{${\textsf{\#id}}$}} (2,1.75);
    \path (1,0.25) edge node[above] {\footnotesize{${1/\textsf{\#d}}$}} (4,0.25);

@@

    \draw (4,-0.25) -- (5,-0.25) -- (5,1.75) -- (4,1.75) -- cycle;
    \node at (4.5,0.75) {\footnotesize{$\texttt{cong}^{l}$}};
    \path (4.5,-0.25) edge node[left] {\footnotesize{$p : \#c \leftrightblackspoon \#d$}} (4.5,-0.75);

    \draw[dashed] (5.8,0.25) -- (5.8,1.25);
    \path [highlight] (5,0.25) edge node[above] {} (6,0.25);
    \path (5,0.25) edge node[above] {\footnotesize{$1/\textsf{\#d}$}} (6,0.25);
    \path [highlight] (5,1.25) edge node[above] {} (6,1.25);
    \path (5,1.25) edge node[above] {\footnotesize{${\textsf{\#d}}$}} (6,1.25);

@@

    \path (9,1.5) edge node[above] {\footnotesize{$1/{\textsf{\#c}}$}} (10,1.5);
  \end{tikzpicture}
  \caption{\label{fig:dual} Dual}
\end{figure}

@@

We also show circuits that construct the \texttt{name} $\ulcorner{p}\urcorner$
and \texttt{coname} $\llcorner{p}\lrcorner$ of $p$, in
figure~\ref{fig:name-coname}.

@@

      \path [highlight] (1,1.75) edge node[above] {} (2,1.75);
      \path (1,1.75) edge node[above] {\footnotesize{$1/\textsf{\#c}$}} (2,1.75);
      \path [highlight] (1,0.25) edge node[above] {} (2,0.25);
      \path (1,0.25) edge node[above] {\footnotesize{${\textsf{\#c}}$}} (2,0.25);
      \draw[dashed] (1.25,1.75) -- (1.25,0.25);

      \draw (2,0) -- (3,0) -- (3,2) -- (2,2) -- cycle;
      \node at (2.5,1) {\footnotesize{$\texttt{cong}^{r}$}};

      \path (3,1.75) edge node[above] {\footnotesize{$1/\textsf{\#c}$}} (4,1.75);
      \path (3,0.25) edge node[above] {\footnotesize{${\textsf{\#d}}$}} (4,0.25);

      \path (2.5,0) edge node[left] {\footnotesize{$p : \#c \leftrightblackspoon \#d$}} (2.5,-0.5);
    \end{tikzpicture}
  \end{subfigure}
  \begin{subfigure}{.5\textwidth}
    \centering
    \begin{tikzpicture}[scale=0.9,every node/.style={scale=0.7},highlight/.style={line width=4pt,orange!70}]

      \path (-1.1,1.75) edge node[above] {\footnotesize{$1/\textsf{\#c}$}} (0,1.75);
      \path (-1.1,0.25) edge node[above] {\footnotesize{${\textsf{\#d}}$}} (0,0.25);

      \draw (0,0) -- (1,0) -- (1,2) -- (0,2) -- cycle;
      \node at (0.5,1) {\footnotesize{$\texttt{cong}^{r}$}};

      \path (0.5,0) edge node[right] {\footnotesize{$p^{-1} : \#d \leftrightblackspoon \#c$}} (0.5,-0.5);

      \path [highlight] (1,1.75) edge node[above] {} (2,1.75);
      \path (1,1.75) edge node[above] {\footnotesize{$1/\textsf{\#c}$}} (2,1.75);
      \path [highlight] (1,0.25) edge node[above] {} (2,0.25);
      \path (1,0.25) edge node[above] {\footnotesize{${\textsf{\#c}}$}} (2,0.25);
      \draw[dashed] (1.75,1.75) -- (1.75,0.25);

@@

      \path (3,1) edge node[above] {\footnotesize{${\textsf{\#id}}$}} (4,1);
    \end{tikzpicture}
  \end{subfigure}
  \caption{\label{fig:name-coname} \texttt{name} and \texttt{coname}}
\end{figure}

>>>>

Compact closed categories have various interesting notions, such as
duals, name and coname, which eventually lead to a certain notion of 
higher order functions.  Here we show that the first three (duals,
name and coname) are definable circuits in $\Pi^/$.

@@

Given a \textrm{/-combinator} $p : \#c \leftrightblackspoon \#d$,
we can define a \emph{dual morphism}
$p^{\ast} : 1/\#d \leftrightblackspoon 1/\#c$. This is shown by building a
reversible circuit in Figure~\ref{fig:dual}.  The highlights that seem to
``escape'' the circuit indicate that \emph{dual} can only be used in a
context with at least one quantifier being present.

\begin{figure}[htb]
  \centering
  \begin{tikzpicture}[scale=0.9,every node/.style={scale=0.7},highlight/.style={line width=4pt,orange!70}]
    \path [highlight] (-1,1) edge node[above] {} (0,1);
    \path (-1,1) edge node[above] {\footnotesize{$1/{\textsf{\#d}}$}} (0,1);

@@

    \path (1,1.75) edge node[above] {\footnotesize{${\textsf{\#id}}$}} (2,1.75);

@@

    \draw (4,0.75) -- (5,0.75) -- (5,1.75) -- (4,1.75) -- cycle;
    \node at (4.5,1.25) {\footnotesize{$p$}};

    \draw[dashed] (5.8,0.25) -- (5.8,1.25);
    \path [highlight] (1,0.25) edge node[above] {} (6,0.25);
    \path (1,0.25) edge node[above] {\footnotesize{${1/\textsf{\#d}}$}} (6,0.25);
    \path [highlight] (5,1.25) edge node[above] {} (6,1.25);
    \path (5,1.25) edge node[above] {\footnotesize{${\textsf{\#d}}$}} (6,1.25);

@@

    \path [highlight] (9,1.5) edge node[above] {} (10,1.5);
    \path (9,1.5) edge node[above] {\footnotesize{$1/{\textsf{\#c}}$}} (10,1.5);
  \end{tikzpicture}
  \caption{\label{fig:dual} Dual}
\end{figure}

@@

We also show circuits that construct the \texttt{name} $\ulcorner{p}\urcorner$
and \texttt{coname} $\llcorner{p}\lrcorner$ of $p$, in
Figure~\ref{fig:name-coname}.  Like with dual, we need to insert a combinator
$p$ \emph{under} a quantifier.

@@

      \path [highlight] (1,1.75) edge node[above] {} (4,1.75);
      \path (1,1.75) edge node[above] {\footnotesize{$1/\textsf{\#c}$}} (4,1.75);

      \path [highlight] (1,0.25) edge node[above] {} (2,0.25);
      \path (1,0.25) edge node[above] {\footnotesize{${\textsf{\#c}}$}} (2,0.25);
      \draw[dashed] (1.25,1.75) -- (1.25,0.25);

      \draw (2,0) -- (3,0) -- (3,0.5) -- (2,0.5) -- cycle;
      \node at (2.5,0.25) {\footnotesize{$p$}};

      \path [highlight] (3,0.25) edge node[above] {} (4,0.25);
      \path (3,0.25) edge node[above] {\footnotesize{${\textsf{\#d}}$}} (4,0.25);

    \end{tikzpicture}
  \end{subfigure}
  \begin{subfigure}{.5\textwidth}
    \centering
    \begin{tikzpicture}[scale=0.9,every node/.style={scale=0.7},highlight/.style={line width=4pt,orange!70}]

      \path [highlight] (-1.1,1.75) edge node[above] {} (2,1.75);
      \path (-1.1,1.75) edge node[above] {\footnotesize{$1/\textsf{\#c}$}} (2,1.75);

      \path [highlight] (-1.1,0.25) edge node[above] {} (0,0.25);
      \path (-1.1,0.25) edge node[above] {\footnotesize{${\textsf{\#d}}$}} (0,0.25);

      \draw (0,0) -- (1,0) -- (1,0.5) -- (0,0.5) -- cycle;
      \node at (0.50,0.25) {\footnotesize{$p^{-1}$}};

      \path [highlight] (1,0.25) edge node[above] {} (2,0.25);
      \path (1,0.25) edge node[above] {\footnotesize{${\textsf{\#c}}$}} (2,0.25);
      \draw[dashed] (1.75,1.75) -- (1.75,0.25);

@@

      \path (3,1) edge node[above] {\footnotesize{${\textsf{\#id}}$}} (4,1);
    \end{tikzpicture}
  \end{subfigure}
  \caption{\label{fig:name-coname} \texttt{name} and \texttt{coname},
  where $p : \#c \leftrightblackspoon \#d$}
\end{figure}
